\documentclass[12pt]{report}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{array}
\usepackage{bm}
\usepackage{bbm}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[left=2.5cm, right=2.5cm, top=2cm, bottom=2.5cm]
{geometry}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[dvipsnames]{xcolor}
\usepackage[hyperfootnotes=false]{hyperref}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{comment}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
	urlcolor=blue,
	citecolor=red,
}
\usepackage{todonotes}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
  \node[shape=circle,draw,inner sep=1pt] (char) {#1};}}
\newcommand{\indep}{\perp \!\!\! \perp}
\bibliographystyle{plain}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{algpseudocode}
% Remove 0's from the sections and subsection titles
\renewcommand{\thesection}{\arabic{section}}
\newcommand{\cbox}[1]{\raisebox{\depth}{\fcolorbox{black}{#1}{\null}}}
\newcommand{\BT}[1]{\todo[color=orange!40]{BT: #1}}
\newcommand{\iBT}[1]{\todo[inline,color=orange!40]{BT: #1}}
\def\mbE{\mathbb{E}}


\title{Why use conditional permutation importance to control the error rate}

\begin{document}

\maketitle

We use the setting of \cite{miPermutationbasedIdentificationImportant2021}.
%
We want to establish the necessity to rely on a conditional importance sampling.

We consider formula (8) in \cite{miPermutationbasedIdentificationImportant2021}, that represents the actual formula used for importance analysis.
%
It can be rewritten:
\[ \hat{M}_j^{(P)} = \frac{1}{N} \left( \| Y-\hat{\mu}(X^{(j)}) \|^2 - \| Y-\hat{\mu}(X) \|^2 \right)
\]
Which yields:
\[
\hat{M}_j^{(P)} = \frac{1}{N} \left( \| \mu(X)-\hat{\mu}(X^{(j)}) \|^2 - \| \mu(X)-\hat{\mu}(X)  \|^2 + 2 \varepsilon (\hat{\mu}(X)-\hat{\mu}(X^{(j)}))  \right)
\]

In \cite{miPermutationbasedIdentificationImportant2021}, it is argued that this vanishes when $N \rightarrow \infty$.
%
But some of the terms converge to 0 too slowly in the presence of correlations, leading to non-type-1 error control.
%
Let us consider a linear model $\mu(X) = Xw$, that one case immediately rewrite $\mu(X) = X_jw_j + X_{-j}w_{-j}$, in the presence of correlation with in $X$, ie $X_j = \alpha X_{-j} w_{-j} + \delta$, where $\alpha$ is a scalar and $\delta$ a noise vector.
%
We further consider that $X_j$ is a null feature ($w_j=0$).
%


\[
\hat{M}_j^{(P)} = \frac{1}{N} \left( 2  \langle (X_j - X_j^{perm}) \hat{w}_j, X_{-j} (w_{-j} - \hat{w}_{-j} + \varepsilon) \rangle \right)
\]
where
$\langle X_j^{perm}, X_{-j} (w_{-j} - \hat{w}_{-j} \rangle = O(\sqrt{N})$ because the permutation destroys existing correlation, and
$\langle (X_j - X_j^{perm}) w_{-j}, \varepsilon \rangle = O(\sqrt{N})$, because this is a noise term.
%
By contrast,
$\langle (X_j, X_{-j} (w_{-j} - \hat{w}_{-j}) \rangle = O(N)$ due to the the correlation between $X_j$ and $X_{-j}$.
This means that there remains a non-vanishing terms when $N \rightarrow \infty$.

Alternatively, if we can generate $X^{(j)}$ conditionally, then
$\langle (X_j - X_j^{perm}) \hat{w}_j, X_{-j} (w_{-j} - \hat{w}_{-j} = O(\sqrt{N}) \rangle$, because the rightmost term becomes uncorrelated to the right one.

\begin{comment}
  We obtain
\[
\mbE(\hat{M}_j^{(P)}) = \frac{1}{N} \left( 2 \alpha \hat{w}_j^2  \langle X_{-j} w_{-j}, X_{-j} (w_{-j} - \hat{w}_{-j}) \rangle \right)
\]

This quantity only slowly converges to $0$ when $N$ increases, because $\hat{w}_j$ converges to $0$ but is
non-zero for finite $N$.
%
This explains why unconditional importance tends to wrongly select
such variables, leading to an uncontrolled error rate. This situation worsens
when $\alpha$ increases.
%
By contrast, if we can generate $X^{(j)}$ conditionally, then the above becomes
\[
\mbE(\hat{M}_j^{(P)}) = \frac{1}{N} \left( 2 \hat{w}_j^2  \mbE(\langle \delta - \delta^{perm}, X_{-j} (w_{-j} - \hat{w}_{-j}) \rangle) \right) = 0
\]

Indeed, $X^{(j)}$ retains the dependency between $X_j'$ and $X_{-j}$ that exists in $X$
and the above term vanishes, leading to an expectancy-0 importance
score and correct inference.
\end{comment}


\bibliography{MyLibrary_dnn}

\end{document}
